---
title: "Control flow"
author: "Wing-Ho Ko"
output: html_notebook
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(stringr)
```


## For loops

A simple `for` loop

```{r}
total <- 0

for (i in 1:500){
  total <- total + i
}

print(total)
```

A for loop that maps a list to a vectors to a vector of sums

```{r}
list_of_vecs <- list(2, c(3, 4), c(2, 5, 6), 4)
out = c()

for (x in list_of_vecs){
  out = c(out, sum(x))
}

print(out)
```

Loop through multiple vectors using `seq_along()`

```{r}

price <- c(18, 42, 51, 38, 25)
quantity <- c(2, 3, 7, 6, 5)

revenue <- 0

for (i in seq_along(price)){
  revenue <- revenue + price[[i]] * quantity[[i]]
}

print(revenue)
```

Using `unique()` to count occurrence:

```{r}
labels  <- c("a", "b", "c", "c", "a", "c", "b")

unique_labels <- unique(labels)
counts <- c()

for (x in unique_labels){
  counts[x] <- sum(labels == x)
}

print(counts)
```

## Conditional statements

Use conditional statement in a `for` loop for filtering

```{r}
values <- c(1, 8, 5, 4, 7, 2)
out = c()


for (val in values){
  if (val %% 2 == 0){
    out = c(out, val)
  }
}

print(out)
```

Including "else if" and "else" in a conditional statements

```{r}

# simulate the output of, e.g., list.files()

filenames <- c("data_2024-06.csv", "data_2025-02.csv", "info.txt", "data_2024-11.csv", "out.zip")

# loop through file names

for (name in filenames){

  ext = str_sub(name, -4, -1)

  if (tolower(ext)==".csv"){
    print(paste0("Data file '", name, "' found"))
  } else if (tolower(ext)==".txt"){
    print(paste0("Metadata file '", name, "' found"))
  } else {
    print(paste0("Additional file '", name, "' found"))
  }
}
```

Compound condition:

```{r}
values <- c(2, 3, NA, 7, -6, 5, NA)

for (val in values){
  if (!is.na(val) && (val < 0 || val > 5)){
    print(val)
  }
}
```

Using `next` for filtering:

```{r}
values <- c(7, 2, NA, 5, 9, 8)

out <- c()

for (val in values){
  
  if (is.na(val) || val %% 2 == 1){
    next
  }
  out = c(out, val)

}

print(out)

```

Using `break` to pick the first element

```{r}
values <- c(-7, -2, 4, 2, NA, -3)

for (val in values){
  if (val > 0){
    break
  }
}

print(val)
```

Put the previous example in a function

```{r}
first_positive <- function(x){
  
  for (val in x){
    if (!is.na(val) && val > 0){
      return(val)
    }
  }
  return(NA)
}

first_positive(c(-4, -7, -1, NA, 5, 3, 4))
```

Return `NA` if the loop finishes without return

```{r}
first_positive(c(-4, -2, -7, -3))
```

## Vectorized conditionals and mapping

Simulate the `abs()` function on vectors using `ifelse()`

```{r}
x <- c(5, -7, -9, 2, -4, NA, 8, 0)

ifelse(x > 0, x, -x)
```

Apply the `seq()` function to a vector to generate a list of vectors

```{r}
lapply(c(2, 4, 5, 3), FUN=seq)
```

Apply both `mean()` and `sd()` function simultaneously on a list of vectors:

```{r}
values = list(c(2, 3), c(-5, 4, 2), c(7, 1, 0, -2))

vapply(values, function(x){c(mean(x), sd(x))}, c(0, 0))
```
