---
title: "Untitled"
author: "Wing-Ho Ko"
output: html_notebook
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(palmerpenguins)
```


## Prelude: dataframe

Define a simple dataframe

```{r}
dframe = data.frame(
  x = c(1, 3, 5),
  y = c(2.5, 7.1, 3.4),
  z = c("here", "there", "where")
)
```

The class of a dataframe is `data.frame`

```{r}
class(dframe)
```

Once unclassed, the dataframe is exposed as a list

```{r}
class(dframe) <- NULL
str(dframe)
```

The elements of the list are columns of the dataframe

```{r}
dframe$x
```

## Subsetting and extending vectors and lists

### Subsetting a vector

Define a simple vector

```{r}
x <- c(-1, 0, 1, 3, 5, 7, 9)
```

Subset by positive integers:

```{r}
x[c(2,3,5)]
```

Subset by negative integers:

```{r}
x[c(-1, -5)]
```

Subset by logical vectors:

```{r}
x[c(TRUE, TRUE, FALSE, FALSE, FALSE, FALSE, TRUE)]
```

Subset the entire vector

```{r}
x[]
```

### Subsetting a list

```{r}
y <- list(-1, 0, 1, 3, 5, 7, 9)
```

Subset by positive integers

```{r}
y[c(5, 7)]
```

Subset by negative integers

```{r}
y[c(-2, -4, -6)]
```

Even with only a single index, the result is a list

```{r}
y[2]
```

Use [[]] to extract the inner object

```{r}
x[[2]]
```

### Assigning new values to elements of vector and list

```{r}
x <- c(-1, 0, 1, 3, 5, 7, 9)
y <- list(-1, 0, 1, 3, 5, 7, 9)
```


Assignment to elements of vector using [] and <-

```{r}
x[2:3] <- c(15, 17)
print(x)
```

Assignment to elements of list using [] and <-

```{r}
y[2:4] <- c(1, 3, 6)
print(y)
```

Assignment to element of list using [[]] and <-

```{r}
y[[5]] <- c(1, 3, 6)
print(y)
```

### Extending a vector

```{r}
x <- c(1, 2, 3)
x <- c(x, c(-2, -4, -5), 0)
print(x)
```

### Extending a list

```{r}
x <- list("a", 2L, 7.4)
```

By default, the second argument of `append()` will be unpacked:

```{r}
y <- c(3L, 1L, 4L)
append(x, y)
```

Avoid unpacking by wrapping `y` inside a list

```{r}
append(x, list(y))
```

## Attributes 

```{r}
dframe = data.frame(
  x = c(1, 3, 5),
  y = c(2.5, 7.1, 3.4),
  z = c("here", "there", "where")
)
```

List the attributes of a simple dataframe

```{r}
attributes(dframe)
```

Assign a new attribute to the dataframe

```{r}
attr(dframe, "ref") <- "Unknown"
```

Check that the attribute is indeed assigned

```{r}
attributes(dframe)
```

Check the names and class of the dataframe

```{r}
names(dframe)
```

```{r}
class(dframe)
```

Turn a dataframe back to the underlying list by removing its class

```{r}
dframe2 <- dframe
class(dframe2) <- NULL
```

```{r}
dframe2
```

Perform an ANOVA on the dependence of body mass on species from the penguin dataset

```{r}
pg_aov <- aov(body_mass_g ~ species, penguins)
```

The object is of class "aov", which is a subclass of "lm"

```{r}
class(pg_aov)
```

The `summary()` call on the aov object is dispatched to `summary.aov()`

```{r}
summary(pg_aov)
```

A different summary is printed out if we remove the "aov" string from the class attribute

```{r}
pg_aov2 <- pg_aov
class(pg_aov2) <- "lm"
summary(pg_aov2)
```

### Named vectors and named lists

Create a named vector by assigning to `names()` after the vector is created

```{r}
z <- c(1, 3, 7)
names(z) <- c("a", "b", "c")
print(z)
```

Create a named vector using key=value syntax

```{r}
x <- c(a = 1, b = 3, c = 7)
print(x)
```

Create a named list using key=value syntax

```{r}
y <- list(a=3, b="this", c=3.14)
print(y)
```

Subsetting named vector

```{r}
x[c("a", "c")]
```

```{r}
x[]
```


Subsetting a named list

```{r}
y[["c"]]
```

Assign an element to y with non-syntactical name

```{r}
y["1 bad name"] <- 6.2
```

Extract the element using back-tick

```{r}
y$`1 bad name`
```

## Inspecting an object with str()

Construct an `lm` object:

```{r}
penguin_lm <- lm(body_mass_g ~ flipper_length_mm, penguins)
```

Inspecting the structure of the lm object

```{r}
str(penguin_lm)
```

## Operations on dataframes and matrices

### Subsetting operations on dataframe

```{r}
dframe = data.frame(
  x = c(1, 3, 5),
  y = c(2.5, 7.1, 3.4),
  z = c("here", "there", "where")
)
```

Subsetting using 1-argument form of []:

```{r}
dframe[c("x", "z")]
```

Subsetting using 2-argument form of []:

```{r}
dframe[1:2, ]
```

Subsetting using 1-argument form of [[]]:

```{r}
dframe[["z"]]
```

Subsetting using 2-argument form of [[]]:

```{r}
dframe[[2, "y"]]
```

### Extending a dataframe

Adding a new column to a dataframe

```{r}
dframe <- data.frame(
  x = 1:5,
  y = seq(2, 10, 2)
)

z <- c(-2, -1, 0, 1, 2)

dframe <- cbind(dframe, z)
dframe
```

Combining two dataframes row-wise

```{r}
dframe1 <- data.frame(
  x = c(1, 2),
  y = c(2, 4)
)

dframe2 <- data.frame(
  y = c(-2, 2),
  x = c(7, -5)
)

rbind(dframe1, dframe2)
```

### Creating a matrix

Creating a matrix

```{r}
mat0 <- matrix(1:9, nrow=3, ncol=3)
print(mat0)
```

Creating the matrix with row-first convention

```{r}
mat <- matrix(1:9, nrow=3, ncol=3, byrow=TRUE)
print(mat)
```

### Assign row names and column names to matrix

```{r}
rownames(mat) <- c("Good", "Neutral", "Evil")
colnames(mat) <- c("Lawful", "Neutral", "Chaos")
print(mat)
```

### Subsetting a matrix

Define a simple matrix

```{r}
mat <- matrix(1:9, nrow=3, ncol=3, byrow=TRUE)
```

Subsetting the matrix

```{r}
mat[1:2, 2:3]
```

Subsetting the underlying vector

```{r}
mat[2:5]
```

### Extending a matrix

```{r}
mat <- cbind(mat, c(-1, 0, 1))
mat <- rbind(mat, c(-4, -3, -2, -1))
print(mat)
```

