---
title: "functions demo"
author: "Wing-Ho Ko"
output: html_notebook
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Writing functions

### Basic example

A simple function that takes an input vector and create a mean-subtracted version of it

```{r}
mean_subtracted <- function(x){
  x - mean(x)
}
```

Apply the function on a vector and assign the return to a new vector

```{r}
x <- c(1, 8, 5, 7, 9)
y <- mean_subtracted(x)
print(y)
```

### Explicit return

Make a function return early (before all lines are executed) using `return()`

```{r}
standardize_bad <- function(x){
  return(x - mean(x))
  x / sd(x) # useless line that never got executed
}
```

```{r}
standardize_bad(c(-1, 0, 1))
```

### Scoping and copy-on-write behaviors

A function cannot modify a global variable

```{r}
x <- 5  # global x
y <- -1 # global y

swap <- function(){

  # a function that, alas, does nothing
  
  x <- 2  # local x
  y <- 10 # local y
}

print(x)
print(y)
```

### Function with default argument

Define the p-norm of a vector (p = 2 is the usual Euclidean norm)

```{r}
vec_norm <- function(x, p = 2){
  return(sum(x^p)^(1/p))
}
```

Use the p-norm function with its default value

```{r}
vec_norm(c(3, 4))
```

Override the default value

```{r}
vec_norm(c(3, 4), 1)
```

### Customize the behavior of a generic function on a new class

```{r}
print.range <- function(x){
  cat(paste0(min(x), " - ", max(x)))
}

x <- c(1, 4)
class(x) <- c("range", class(x))

print(x)
```
